---
title: "datasets"
author: "John Morse"
date: "2/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include = FALSE}
library(readxl)
library(janitor)
library(tidyverse)
library(gt)
library(rvest)
library(reprex)
library(fivethirtyeight)
```

My work so far: My data comes from a variety of sources that includes the CDC,
academic research, and a pharmaceutical company.

I have spent most of my time cleaning the data and ensuring that all of the
disparate graphs and tables are in proper form so that I can run analysis. My
work so far has looked like this:


1. Gather data and store on my computer 

2. Create GitHub repo which can be found here:
https://github.com/jrmorse/final_project_datasets

3. Load data into my .rmd


4. scrape data from a website into my RMD. Here I had to spend a lot time
ensuring that the format was correct. This meant getting rid of the costs for
diseases that I wasn’t interested in, converting characters to integers, and
segmenting between direct and indirect medical costs.


5.  I made a function to run through a folder in my repo and to automatically,
pull, clean, and combine data referring to PrEP usage in the U.S.


6. I made a quick bar graph to show the increase of usage between men and women
between 2012-2018.


7. Then I read in, cleaned, and reformatted all CDC data for HIV and other
relevant STIs.

*Please see below for the tables I have created*

```{r}
# This includes the direct medical costs associated with key STIs in the U.S.. I
# will be looking at the costs of HIV, chlamydia, gonorrhea, and I think
# syphilis as well. The prices listed are in 2006 USD, so I will create another
# column that calculates the 2019 USD cost for each infection.I have used the
# pivot long function to create a new column relating to sex. The data should be
# clean and I have used the slice function to get rid of unwanted rows. 

table_costs_messy <- read_html("https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2426671/")%>% 
  html_node("table") %>% 
  html_table(header = TRUE) %>% 
  clean_names() %>% 
  slice(-1) %>% 
  slice(4:17)

# This is where I have made sex a new column so that I can better compare costs
# of illness for men and women. Additionally I have created a column to adjust
# 2006 USD to that of 2020 USD. The rate of inflation between the years was
# 131%.

costs_clean <- table_costs_messy %>% 
  rename(xF = value_applied, xM = x) %>% 
  pivot_longer(
    cols = starts_with("x"),
    names_to = "sex",
    names_prefix = "x",
    values_to = "cost",
    values_drop_na = FALSE
  ) %>% 
# I did this line below to get rid of extra text in the parameter column.
  mutate_all(~ gsub(x =., pattern = "\\*| \\[5]| \\[6]| \\[1,64,65]| \\[85]", "")) %>% 

  mutate(cost_2006 = parse_number(cost)) %>% 
  mutate(cost_2020 = (cost_2006 * 1.31)) %>% 
  select("parameter", "sex", "cost_2020")

costs_clean

# This table does not differentiate between indirect and direct costs as the
# dataset has indirect costs listed as a separate that is embedded in the data.
# In the following r blocks I will separate out the indirectand direct costs so
# that I can be more specific.
```
```{r}
# DIRECT COSTS FOR HIV & STIS

# This is the data that is just focused on direct medical costs.

direct_costs <-
  costs_clean %>% 
  slice(1:10)
colnames(direct_costs)[3] <- "direct_cost_2020"

direct_costs
```


```{r}
# INDIRECT COSTS HIV & STIS

# This is the data that is concerned with the indirect costs of HIV

indirect_costs_hiv <- costs_clean %>% 
  slice(13:14)

# This is the indirect costs of other relevant STIs

indirect_costs_sti <- costs_clean %>% 
  slice(21:28) 

# Here I'll join these two tables together with a bind function.
indirect_costs = rbind(indirect_costs_hiv, indirect_costs_sti)

# change the title of the cost column to be more specific
colnames(indirect_costs)[3] <- "indirect_cost_2020"

indirect_costs

```
```{r}
# INDIRECT & DIRECT COSTS

complete_costs <- direct_costs %>% 
  full_join(indirect_costs, direct_costs, by = c("parameter"= "parameter","sex"="sex")) %>% 
  mutate(total_cost = direct_cost_2020 + indirect_cost_2020)

complete_costs
```


```{r, include = FALSE}
# PREP DATA

# Empty data set where I will store the files once they'e loaded in.
prep_data_1 <- data.frame()

# Listing out all the files in this particular folder. All of this data relates
# to PrEP usage in the U.S..
list <- list.files("Data_Sets/PrEP_Data")

# Here I have the actual function which is picking up on every document that
# ends in ".xlsx" in the "PrEP_Data" folder. The function reads in all of the
# data, and then spits each one into the empty dataframe that I have created.
# Each row is appended onto the preceding year's data.

listxlsx <- dir(pattern = "*.xlsx")
for (i in 1:length(list)){
  print(list[i])
  temp_data <- read_xlsx(paste("Data_Sets/PrEP_Data/",
                               list[i], sep = ""),
                         skip = 1) %>% 
    clean_names()
  
  prep_data_1 = rbind(prep_data_1, temp_data)
}

prep_data_1

```
```{r}
# PrEP growth rates for men and women

# data for only male users
prep_m <- prep_data_1 %>% 
  select(year, male_pr_ep_users, male_pr_ep_rate) %>%
  mutate(sex = "Male")
colnames(prep_m)[2:3] <- c("pr_ep_users", "pr_ep_rate")
prep_m

# data for only female users
prep_f <- prep_data_1 %>% 
  select(year, female_pr_ep_users, female_pr_ep_rate) %>% 
  mutate(sex = "Female")
colnames(prep_f)[2:3] <- c("pr_ep_users", "pr_ep_rate")
prep_f

# Here I'll use the bind function to list the information ontop of eachother,
prep_gender = rbind(prep_f, prep_m)
prep_gender

ggplot(data = prep_gender, aes(x = year, y = pr_ep_users, fill = sex)) +
  geom_col(position = "dodge") +
  labs(title="PrEP Users by Gender", subtitle = "2012-2018") +
  theme_classic() +
  scale_x_continuous(
  breaks = seq(2012,2018,1),
  label = c("2012", "2013", "2014", "2015", "2016","2017","2018")) +
  scale_y_continuous()

```

```{r}
# CHLAMYDIA AND GONORRHEA ONLY
# Here I read in all the CDC data for Chlamydia and Gonorrhea rates from 2000 to
# 2017. I still should rename the columns so that it is clear which values are
# for which variable when I join the datasets. Also, for testing pursposes, I
# have set this to only return the first ten rows as it is a large dataset.


chlamydia_gonorrhea <-read_xlsx("Data_Sets/CDC_Data/ATLAS_CDC_Chlamydia&Gonorrhea.xlsx",
  skip = 5) %>% 
  clean_names() 
chlamydia_gonorrhea[chlamydia_gonorrhea == "Data not available"] <-NA

chlamydia_gonorrhea <- chlamydia_gonorrhea %>%
  filter(!is.na(cases))
chlamydia_gonorrhea

```

```{r HIV DIAGNOSIS ONLY}
# Here I read in all the CDC data for HIV/AIDS rates for females from 2008 to
# 2017. I have added an extra column to list what the sex is.

hiv_aids_f <- read_xlsx("Data_Sets/CDC_Data/Atlas_CDC_HIV:AIDS_Female.xlsx",
  skip = 8) %>% 
  clean_names() %>% 
  mutate(sex = "F")
hiv_aids_f[hiv_aids_f == "Data not available"] <-NA

# Here I read in all the CDC data for HIV/AIDS rates for males from 2008 to
# 2017. I have added an extra column to list what the sex is.

hiv_aids_m <- read_xlsx("Data_Sets/CDC_Data/AtlasPlus_CDC_Male_HIV:AIDS.xlsx",
  skip = 8) %>% 
  clean_names() %>% 
  mutate(sex = "M")
hiv_aids_m[hiv_aids_m == "Data not available"] <-NA

# Here I merged the two HIV/AIDS datasets as the variables were the same for
# both the female and male tables. For testing purposes, I am only showing the
# first 10 rows.

# I filtered for indicator of HIV diagnoses because if I do not, it lumps prevelance and diagnoses together in a way that is misleading for the information I want to communicate.

hiv_aids_all <- rbind(hiv_aids_f, hiv_aids_m) %>% 
  filter(indicator == "HIV diagnoses")

hiv_aids_all_test <- hiv_aids_all %>%
  slice(1:10) %>% 
  gt()

hiv_aids_all_test
```

```{r HIV Cases by year Graph}
hiv_aids_year <- hiv_aids_all %>% 
  group_by(year) %>%
  summarize(total_cases = sum(cases))

 ggplot(hiv_aids_year, aes(x = year, y = total_cases)) +
   geom_point() +
   geom_text(aes(x = year, y = total_cases, label = total_cases),vjust = -1, nudge_y = .5) +
   labs(title="HIV Diagnoses by Year", subtitle = "2007-2017") +
   theme_classic() +
   scale_x_continuous(
     name = "Year",
     breaks = seq(2007,2017,1),
     label = c("2007", "2008", "2009", "2010", "2011","2012","2013","2014","2015","2016","2017")) +
   scale_y_continuous(
     name = "New Diagnoses",
     limits=c(35000, 50000))

```

```{r HIV Diagnoses by Gender Graph}

ggplot(data = hiv_aids_all, aes(x = year, y = cases, fill = sex)) +
  geom_col(position = "dodge") +
  labs(title="HIV Diagnoses by Gender", subtitle = "2007-2017") +
  theme_classic() +
  scale_x_continuous(
    name = "Year",
    breaks = seq(2007,2017,1),
    label = c("2007", "2008", "2009", "2010", "2011","2012","2013","2014","2015","2016","2017")) +
  scale_y_continuous(
    name = "New Diagnoses"
  )
```

```{r HIV Diagnoses Transmission Category Tibble}
hiv_aids_transmission <- hiv_aids_all %>% 
  group_by(transmission_category, year, sex) %>%
  summarize(total_cases = sum(cases))
hiv_aids_transmission
```

```{r HIV Diagnosis by Category Graph}
ggplot(data = hiv_aids_transmission, aes(x = year, y = total_cases, fill = transmission_category)) +
  geom_col(position = "dodge") +
  labs(title="HIV Diagnosis by Category", subtitle = "2007-2017") +
  theme_classic() +
  scale_x_continuous(
    name = "Year",
    breaks = seq(2007,2017,1),
    label = c("2007", "2008", "2009", "2010", "2011","2012","2013","2014","2015","2016","2017")) +
  scale_y_continuous(
    name = "New Diagnoses"
  )
```

```{r Male to male sexual contact}
hiv_aids_maletomalesexualcontact <- hiv_aids_all %>% 
  filter(transmission_category == "Male-to-male sexual contact" & race_ethnicity == c("Black/African American", "White", "Hispanic/Latino")) %>%
  group_by(year, race_ethnicity) %>%
  summarize(total_cases = sum(cases))
hiv_aids_maletomalesexualcontact

ggplot(data = hiv_aids_maletomalesexualcontact, aes(x = year, y = total_cases, fill = race_ethnicity)) +
  geom_col(position = "dodge") +
  labs(title="HIV Diagnosis Male-to-male Sexual Contact", subtitle = "2007-2017") +
  theme_classic() +
  scale_x_continuous(
    name = "Year",
    breaks = seq(2007,2017,1),
    label = c("2007", "2008", "2009", "2010", "2011","2012","2013","2014","2015","2016","2017")) +
  scale_y_continuous(
    name = "New Diagnoses"
  )
```

```{r}
# SYPHILIS ONLY
# Here I read in all the CDC data for both early latent and congential syphilis
# rates from 2008 to 2017. I still should rename the columns so that it is clear
# which values are for which variable when I join the datasets.

syphilis <- read_xlsx("Data_Sets/CDC_Data/Atlas_Congenital&EarlyLatentSyphilis.xlsx",
  skip = 7) %>% 
  clean_names()
syphilis[syphilis == "Data not available"] <-NA

syphilis <- syphilis %>% 
  gt()

syphilis
```

